Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-04-18 paper-correlations.tex
Clearer description of algorithms

\documentclass[reprint,amsmath,amssymb,aps,pre]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

@@

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$:
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot zeros(length(a))$ \Comment{$\odot$
      denotes concatenation and $zeros(n)$ returns an array with $n$ zeros.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets subseq(a, shift)$
      \Comment{$subseq$ returns subsequence of $a$ from the $shift$-th element
        to the end.}
      \State $success[shift] \gets sum(map(\chi, a, a_2))$
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets clusterize(a_1)$
  \Comment{$clusterize(x)$ selects clusters from the array
    $x$. Each cluster gets its own unique integer identificator. The
    identificator $0$ is reserved for the void phase.}
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

>>>>

\documentclass[reprint,amsmath,amssymb,aps,pre,nofootinbib]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

@@

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$: \footnote{From
here on $\odot$ denotes concatenation and $:$ denotes consing (joining an
element in front of a list).}
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot [0 \dots 0]$ \Comment{Pad with zeros to twice the
      length of $a$.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets a[shift \dots]$
      \Comment{Take elements with index bigger than $shift$.}
      \State $success[shift] \gets sum(\chi(a, a_2))$
      \Comment{Apply $\chi$ element-wise and take a sum.}
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

For computation of cluster function we use \code{label\_components} function
from \code{Image.jl} package which extracts connected comonents (clusters) from the
binary array. Each cluster gets its own unique integer identificator. The
identificator $0$ is reserved for the void phase.

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets label\_components(a_1)$
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}
